---
title: "QMF Final Project"
author: "Ajay Dugar"
date: "11/13/2020"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(signal)
library(oce)
library(latex2exp)
library(actuar)
```


## Generate a Poisson process

First, we generate a sequence of independent exponential random variables, with $\lambda$ as the parameter, giving us a CDF of $P[\tau_i \geq y] = e^{-\lambda y}$. Allow $T_n = \sum_{i=1}^n \tau_i$. Using this sequence, $T_n$, we can generate a Poisson process, $N_t$, where $N_t = \sum_{n\geq 1} 1_{t\geq T_N}$

```{r}
set.seed(6083)
n = 20
lambda = 2
tau = rexp(n = n, rate = lambda) #generate independent exponential random variables
T_n = cumsum(tau)
time_steps = 10000
tmax = max(T_n)+1
time = seq(from = 0, to = tmax, length.out = time_steps)
N_t = time
for (i in 1:time_steps){
  N_t[i] = sum(N_t[i] > T_n)
}
plot(time, N_t, ylab = TeX("N_t"), xlab = TeX("t"), type = "l", main =  TeX("Simple Poisson Process, $\\lambda = 2, T_n = 20"))
```

The behavior of this Poisson process is right-continuous with left-sided limits (RCLL), with jumps of 1 at the jump times $T_i$, with the $T_{i+1} - T_{i}$ being exponentially distributed. This gives a probability distribution of the Poisson process for a given time, $t$, and a given rate of occurrence, $\lambda$, of $P[N_t = n] = e^{-\lambda t} \frac{(\lambda t)^n}{n!}$.


$$N_t \sim Pois(k\lambda)$$

This Poisson process is more generally a Lévy process, satisfying the following conditions:
1) Increments are independent, i.e. $E[X_t - X_s|t,s] = E[X_t - X_s], t \geq s$
2) Increments are stationary, specifically $X_t-X_s \sim Poisson(\lambda (t-s)), t \geq s$
3) The process continuous: $lim_{h \rightarrow 0} P(|X_{t+h} - X_t| > \epsilon) = 0; \forall \space \epsilon, \forall t$

Since jump-diffusion processes don't necessarily have a clean closed form distribution function, we can use the characteristic function instead. For a given random variable X, the characteristic function is:

$$\Phi_t(u) = E[e^{iu X}]$$

Specifically for the Poisson process, the characteristic function is:
$$E[e^{iuN_t}] = e^{\lambda t (e^{i u} - 1)}$$
Since we are dealing with the simulation of financial assets, the jump size should not be a singular value. This leads to a compound Poisson process where the distribution of times between jumps remains exponential, but the values of the jumps will have a different distribution. Allowing $N_t$ to remain a Poisson process with parameter $\lambda$, but now let $\{Y_i\}_{i \geq 1}$ be a sequence of independent random variables with distribution function $f(x)$. Then 
$$X_t = \sum_{i=1}^{N_t} Y_i$$

gives a compound Poisson process, $X_t$. The distribution function is not explicitly known for a given time $t$, but the characteristic function has the form:

$$E[e^{iuX_t}|N_t ] = e^{t \lambda \int_\mathbb{R} e^{iux}-1 f(dx)}$$
Simulating this compound process, $X_T$, where you have $n$ independent jumps, i.e. $N_T = n$, on $t \in [0,T]$ requires 3 steps:
1) Simulating $N_T$ from a Poisson distribution with parameter $\lambda T$
2) Simulating $n$ uniform random variables $\{U_i\}_{i=1}^{n}$ on $[0,T]$
3) Simulating $n$ independent variables $\{Y_i\}_{i=1}^{n}$ with distribution $f$

Now we can generate the process:
$$X_t = \sum_{i=1}^{n} Y_i 1_{U_i \leq t}$$

```{r}
set.seed(6083)

# In this case, our jump size will be Gaussian, N(0,1)
Y_i = rnorm(n = length(T_n), mean = 0, sd = 1)

X_t = time
for (i in 1:time_steps){
  X_t[i] = sum((X_t[i]>T_n)*Y_i)
}

plot(time, X_t, ylab = TeX("X_t"), xlab = TeX("t"), type = "l", main =  TeX("Compound Poisson Process, $\\lambda = 2, T_n = 20"))
```

Now the next step is to generate a jump-diffusion process. This will be accomplished by combining 2 different stochastic processes:
1) The Compound Poisson Process $\sum_{i=1}^{N_t} Y_i 1_{U_i \leq t}$
2) Brownian motion with drift $\mu t + \sigma B_t$

```{r}
set.seed(6083)

sigma = 0.1
mu = 0.05
X2_t = X_t + sigma*rnorm(length(time)) + mu*time

plot(time, X2_t, ylab = TeX("X_t"), xlab = TeX("t"), type = "l", main =  TeX("Jump-Diffusion Process, $\\lambda = 2, T_n = 20, \\sigma = 0.25, \\mu = 0.05"))
```

Now we can simulate an exponential Levy model which may be commonly referred to as the Merton Model:

$$S_t = S_0 e^{X_t}$$

where $X_t$ is the jump-diffusion process from above.

```{r}
S_0 = 10

S_t = S_0*exp(X2_t)

plot(time, S_t, ylab = TeX("S_t"), xlab = TeX("t"), type = "l", main =  TeX("Merton Model"))
```

This form can be written as a generalization of the basic Black-Scholes model:

$$\frac{dS_t}{S_{t-}} = (\mu+r)dt+\sigma dB_t + dJ_t$$
With
$$dJ_i = e^{Y_i}-1$$

Let us consider the payoff of a geometric Asian call option with this model. The payoff is as following:

$$C(K) = max(A(0, T) - K, 0)$$
Where:

$$A(0, T) = exp(\frac{1}{T} \int_{0}^{T} ln(S_t) dt) = exp(\frac{ln(S_0)}{T} \int_{0}^{T} (\mu+r) t + \sigma B_t + \sum_{i=1}^{N_t} Y_i \space dt) $$

Let us simulate the Merton Model for the Asian option prices:

```{r}
spot_prices = seq(90, 110, length.out = 200)
K = 100
r = 0.05
sigma = 0.02
T = 1

sigma_tilde = 0.2
lambda = 3

S_0 = 100
```

```{r}
set.seed(6083)

asian_prices = c()

for (i in spot_prices){
  
  simulated_values = c()
  
  for (j in seq(1, 10)){
    
    n = 252
    tau = rexp(n = n, rate = lambda) 
    T_n = cumsum(tau)
    time_steps = 1000
    tmax = max(T_n)+1
    time = seq(from = 0, to = tmax, length.out = time_steps)
    N_t = time
    for (i in 1:time_steps){
      N_t[i] = sum(N_t[i] > T_n)
    }
    
    Y_i = rnorm(n = length(T_n), mean = 0, sd = sigma_tilde)
    
    X_t = time
    for (i in 1:time_steps){
      X_t[i] = sum((X_t[i]>T_n)*Y_i)
    }
    
    X2_t = X_t + rnorm(length(time), sd = sigma)
    
  }
  
  asian_prices = c(X2_t, mean(X2_t))
}
```

Taking a look at one of our reference underlying paths:

```{r}
plot(X2_t, type = "l", ylab = TeX("X_t"), xlab = "t")
```


Let us compare this to a vanilla European call option, with payoff:

$$C(K) = \Phi( \frac{ln(\frac{S_T}{K}) + (r+\frac{\sigma^2}{2})T}{\sigma \sqrt{T}}) - \Phi( \frac{ln(\frac{S_T}{K}) + (r+\frac{\sigma^2}{2})T}{\sigma \sqrt{T}} - \sigma \sqrt{T})Ke^{-rT}$$

```{r}
T = 0.25

d1 = (log(spot_prices/K) + (r + 0.5*sigma^2)*T)/(sigma*sqrt(T))
d2 = d1 - sigma*sqrt(T)
euro_gbm = pnorm(d1)*spot_prices - pnorm(d2)*K*exp(-r*T)

d1 = (log(spot_prices/K) + (r + 0.5*sqrt(sigma^2 + lambda*sigma_tilde^2*T)^2)*T)/(sqrt(sigma^2 + lambda*sigma_tilde^2*T)*sqrt(T))
d2 = d1 - sqrt(sigma^2 + lambda*sigma_tilde^2*T)*sqrt(T)
euro_jd = pnorm(d1)*spot_prices - pnorm(d2)*K*exp(-r*T)

sigma_g = sigma/sqrt(3)
b = 0.5*(r-0.5*sigma_g^2)
d1 = (log(spot_prices/K)+(b+0.5*sigma_g^2)*T)/(sigma_g*sqrt(T))
d2 = d1 - sigma_g*sqrt(T)
asian_gbm = pnorm(d1)*spot_prices*exp((b-r)*T)-K*exp(-r*T)*pnorm(d2)

sigma_g = sqrt(sigma^2 + lambda*sigma_tilde^2*T)/sqrt(3)
b = 0.5*(r-0.5*sigma_g^2)
d1 = (log(spot_prices/K)+(b+0.5*sigma_g^2)*T)/(sigma_g*sqrt(T))
d2 = d1 - sigma_g*sqrt(T)
asian_jd = pnorm(d1)*spot_prices*exp((b-r)*T)-K*exp(-r*T)*pnorm(d2)
  
plot(spot_prices, euro_gbm, type = "l", col = "black", xlim = c(90, 105), ylim = c(0, 10), xlab = "Spot Prices", ylab = "Option Value", main = "Option values 3 months from maturity")
lines(x = spot_prices, y = euro_jd, type = "l", col = "green")
lines(x = spot_prices, y = asian_gbm, type = "l", col = "blue")
lines(x = spot_prices, y = asian_jd, type = "l", col = "red")
lines(x = spot_prices, y = asian_sim, type = "l", col = "cyan")
legend(90, 6, legend=c("European (GBM)", "European (JD)", "Geometric Asian (GBM)", "Geometric Asian (JD)"), col=c("black","green", "blue", "red"), lty=1, cex=0.8)
```

```{r}
T = 1/365

d1 = (log(spot_prices/K) + (r + 0.5*sigma^2)*T)/(sigma*sqrt(T))
d2 = d1 - sigma*sqrt(T)
euro_gbm = pnorm(d1)*spot_prices - pnorm(d2)*K*exp(-r*T)

d1 = (log(spot_prices/K) + (r + 0.5*sqrt(sigma^2 + lambda*sigma_tilde^2*T)^2)*T)/(sqrt(sigma^2 + lambda*sigma_tilde^2*T)*sqrt(T))
d2 = d1 - sqrt(sigma^2 + lambda*sigma_tilde^2*T)*sqrt(T)
euro_jd = pnorm(d1)*spot_prices - pnorm(d2)*K*exp(-r*T)

sigma_g = sigma/sqrt(3)
b = 0.5*(r-0.5*sigma_g^2)
d1 = (log(spot_prices/K)+(b+0.5*sigma_g^2)*T)/(sigma_g*sqrt(T))
d2 = d1 - sigma_g*sqrt(T)

asian_gbm = pnorm(d1)*spot_prices*exp((b-r)*T)-K*exp(-r*T)*pnorm(d2)

sigma_g = sqrt(sigma^2 + lambda*sigma_tilde^2*T)/sqrt(3)
b = 0.5*(r-0.5*sigma_g^2)
d1 = (log(spot_prices/K)+(b+0.5*sigma_g^2)*T)/(sigma_g*sqrt(T))
d2 = d1 - sigma_g*sqrt(T)
asian_jd = pnorm(d1)*spot_prices*exp((b-r)*T)-K*exp(-r*T)*pnorm(d2)
  
plot(spot_prices, euro_gbm, type = "l", col = "black", xlim = c(99.7, 100.1), ylim = c(0, 0.15), xlab = "Spot Prices", ylab = "Option Value", main = "Option values 1 day from maturity")
lines(x = spot_prices, y = euro_jd, type = "l", col = "green")
lines(x = spot_prices, y = asian_gbm, type = "l", col = "blue")
lines(x = spot_prices, y = asian_jd, type = "l", col = "red")
lines(x = spot_prices, y = asian_sim, type = "l", col = "cyan")
legend(99.7, 0.15, legend=c("European (GBM)", "European (JD)", "Geometric Asian (GBM)", "Geometric Asian (JD)"), col=c("black","green", "blue", "red"), lty=1, cex=0.8)
```


# References
R. Merton, Option pricing when underlying stock returns are discontinuous, J. Financial Economics, 3 (1976), pp. 125–144.

P. Carr and D. Madan, Option valuation using the fast Fourier transform, J. Comput. Finance, 2 (1998), pp. 61–73.

R. Cont and P. Tankov, Financial Modelling with Jump Processes,
Chapman & Hall / CRC Press, 2004.

P. Tankov, L´evy Processes in Finance: Inverse Problems and Dependence
Modelling, PhD thesis, Ecole Polytechnique, France, 2004.

Kemna, A.G.Z.; Vorst, A.C.F.; Rotterdam, E.U.; Instituut, Econometrisch (1990), A Pricing Method for Options Based on Average Asset Values

Tankov, Peter & Voltchkova, Ekaterina. (2009). Jump-difiusion models: a practitioner's guide. Banque et Marchés. 